---
title: "Week 9 Live Session"
author: "w203 Instructional Team"
date: "Oct 24, 2016"
output: pdf_document
---
## Comparing Means

The file united\_states\_senate\_2014.csv contains data on the 100 members of the US senate that served in 2014.  We will consider this group to be a sample (for example, from some generative process that creates senators).

```{r}
S = read.csv("united_states_senate_2014.csv")
summary(S)
```

You will be placed in a breakout room and assigned a single question to investigate using this dataset.  Each question involves a comparison of means.



S$protestant = S$Religion == "Protestant"
summary(S$protestant)
summary(S$Campaign.Money.Spent)
t.test(S$Campaign.Money.Spent ~ S$protestant)



In your breakout rooms, examine the data and decide what type of test is most appropriate.  You may select a paired or an unpaired test.  You may also select a parametric or a nonparametric test.  Conduct your test and interpret your results.

Room 1: Is there a difference between the amount of money a senator raises and the amount spent?

Room 2: Do female Democratic senators raise more or less money than female Republican senators?

Room 3: Do protestant Senators spend more or less money than non-protestant senators?


Room 4: Choose your own question to investigate.






## Demonstration of Confidence Intervals

The following exercise is meant to demonstrate what the confidence level in a confidence interval represents.  For this exercise, we will assume a standard normal population distribution and simulate what happens when we draw a sample and compute a confidence interval.

Your task is to complete the following function so that it,

1) simulates and stores n draws from a standard normal distribution
2) based on those draws, computes a valid confidence interval with confidence level $\alpha$.

Your function should return a vector of length 2, containing the lower bound and upper bound of the confidence interval.

```{r}
sim_conf_int = function(n, alpha) {
  # Your code to 
  # 1. simulate n draws from a standard normal dist.
  # 2. compute a confidence interval with confidence level alpha
 return(c(-1,1))  # replace with the interval you compute.
}
```

When your function is complete, you can use the following code to run your function 100 times and plot the results.

```{r}
many_conf_int = function(m, n, alpha) {
  results = NULL
  for(i in 1:m) {
    interval = sim_conf_int(n, alpha)
    results = rbind(results, c(interval[1], interval[2], interval[1]<0 & interval[2]>0))
  }
  resultsdf = data.frame(results)
  names(resultsdf) = c("low", "high", "captured")
  return(resultsdf)
}

n = 20
cints = many_conf_int(100, n, .05)

plot(NULL, type="n",xlim=c(1,100),ylim=c(min(cints$low), max(cints$high)), xlab="Trial",ylab=expression(mu),pch=19)
abline(h = c(0, qt(0.975, n-1)/sqrt(n), qt(0.025, n-1)/sqrt(n)), lty = c(1,2,2), col = "gray")
points(cints$high, col = 2+cints$captured, pch = 20)
points(cints$low, col = 2+cints$captured, pch = 20)
for(i in 1:100)
   {
     lines(c(i,i), c(cints$low[i],cints$high[i]), col = 2+cints$captured[i], pch = 19)
     }
title(expression(paste("Simulation of t-Confidence Intervals for ", mu,
                          " with Sample Size 20")))
legend(0,-.65, legend = c(expression(paste(mu," Captured")),
                             expression(paste(mu," Not Captured"))), fill = c(3,2))

```

