---
title: "Reilly_Adam_HW1"
author: "Adam Reilly"
date: "September 7, 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd ("C:/Users/PC/Desktop/Statistics for DS")
getwd

cars = read.table('cars.csv', sep = ",", header = TRUE, na.strings = "na")
cars
```

1. What are the variables in the file?
```{r}
ls(cars)
```

2. Find the mean, median, minimum, maximum, 1st quartile and 3rd quartile for the mpg variable.
```{r}
mean(cars$mpg)
median(cars$mpg)
min(cars$mpg)
max(cars$mpg)
quantile(cars$mpg)
```

3. Create a histogram of the mpg variable.
```{r}
hist(cars$mpg, breaks = , main = 'Histogram of Miles Per Gallon' )
```

4. What is the standard deviation of mpg variable?
```{r}
sd(cars$mpg)
```

5. What is the variance of mpg variable?
```{r}
var(cars$mpg)
```

6. What is the relationship of the standard deviation to the variance? Why does the standard deviation and variance of the mpg variable differ?

The standard deviation is the square root of the variance. They differ in that standard deviation is on the same scale as the mean, whereas variance is not. Variance is calculated by taking the sum of the square of the difference between every point in a data set and the mean, then dividing that sum by the number of data points.

7. How many data points are there for the cyl variable?
```{r}
cyl_valid = subset(cars, is.na(cars$cyl) == F)
nrow(cyl_valid)
```

8. What is the mean of the cyl variable?
```{r}
mean(cyl_valid$cyl)
```

