---
title: "Adam_Reilly_HW3"
author: "Adam Reilly"
date: "September 15, 2016"
output: pdf_document
---

```{r setup, include=FALSE}
setwd ("C:/Users/PC/Desktop/Statistics for DS")
getwd
knitr::opts_chunk$set(echo = TRUE)
```


##1. Gas Station Analytics 
At a certain gas station, 40% of customers use regular gas (event R), 35% use mid-grade (event M), and 25% use premium (event P). Of the customers that use regular gas, 30% fill their tanks (Event F). Of the customers that use mid-grade gas, 60% fill
their tanks, while of those that use premium, 50% fill their tanks. Assume that each customer is drawn independently from the entire pool of customers.

Setup
\newline
P(R) = 0.4
\newline
P(M) = 0.35
\newline
P(P) = 0.25
\newline
P(R and F) = 0.4 * 0.3 = 0.12
\newline
P(M and F) = 0.35 * 0.6 = 0.21
\newline
P(P and F) = 0.25 * 0.5 = 0.125
\newline
P(F) = 0.12 + 0.21 + 0.125 = .455
\newline

(a) What is the probability that the next customer will request regular gas and fill the tank?
\newline
P(R and F) = 0.12
\newline
(b) What is the probability that the next customer will fill the tank?
\newline
P(F) = 0.455
\newline
(c) Given that the next customer fills the tank, what is the conditional probability that they
use regular gas?
\newline
P(R|F) = P(R and F)/P(F) = 0.12/0.455 = 0.264
\newline

##2. The Toy Bin
In a collection of toys, 1/2 are red, 1/2 are waterproof, and 1/3 are cool. 1/4 are red and
waterproof. 1/6 are red and cool. 1/6 are waterproof and cool. 1/6 are neither red, waterproof,
nor cool. Each toy has an equal chance of being selected.

Setup
\newline
P(Red) = 0.5
\newline
P(Waterproof) = 0.5
\newline
P(Cool) = 0.333 
\newline
P(Red and Waterproof) = 0.25
\newline
Since P(Red|Waterproof) = P(Red and Waterproof)/P(Waterproof) = 0.25/0.5 = 0.5 = P(Red)
\newline
Red and Waterproof are independent
\newline
P(Red and Cool) = 0.167
\newline
Since P(Red|Cool) = P(Red and Cool)/P(Cool) = 0.167/0.333 = 0.5 = P(Red)
\newline
Red and Cool are independent
\newline
P(Waterproof and Cool) = 0.167
\newline
Similar to above, this leads to waterproof and cool being independent.
\newline
P(!Red or !Waterproof or !Cool) = 0.167
\newline
P(Red, Waterproof and Cool) = 0.083
\newline
Since all three variables are independent, we know the P(All Three) = 0.083 (Which can be determined various ways (P(Cool) * P(Red and Waterproof), P(Red) * P(Cool and Waterproof)

(a) Draw an area diagram to represent these events.

```{r pressure, echo=FALSE}
labels = c("Red", "Waterproof", "Cool", "Red and Waterproof", "Red and Cool", "Cool and Waterproof", "Red, Cool and Waterproof", "None")
amounts = c(2, 2, 1, 2, 1, 1, 1, 2)
pie(amounts, labels)
```

(b) What is the probability of getting a red, waterproof, cool toy?
\newline
0.083

(c) You pull out a toy at random and you observe only the color, noting that it is red.
Conditional on just this information, what is the probability that the toy is not cool?
\newline
P(!Cool|Red) = (P(Red)-P(Red and Cool))/P(Red) = 
\newline(0.50-0.167)/0.5 = 
\newline(0.333/0.5) = 
\newline 0.667

(d) Given that a randomly selected toy is red or waterproof, what is the probability that it is
cool?
P(Cool|Red or Waterproof) = P(Cool and Red or Cool and Waterproof)/P(Red or Waterproof) = 
\newline
(P(Cool and Red)+P(Cool and Waterproof)-P(Cool, Red and Waterproof))/(P(Red) + P(Waterproof) - P(Red and Waterproof) =
\newline
(0.167+0.167-0.083)/(0.5 + 0.5 - 0.25) =
\newline
0.25/0.75=
\newline
0.333

##3. On the Overlap of Two Events
Suppose for events A and B, P(A) = 1/2, P(B) = 2/3, but we have no more information about
the events.
\newline

(a) What are the maximum and minimum possible values for P(A intersection B)?
\newline
The maximum value for the intersection assumes that all A is in B, or P(A and B) = 
\newline
0.5
\newline
The minimum value for the intersection assumes that as little of A in possible is in B =  (P(A and B)) = 
\newline
0.167
\newline

(b) What are the maximum and minimum possible values for P(A|B)?
\newline
Given that B is true, the maximum value for P(A) would be if as much of A overlapped with B as possible (or all of it)
\newline
The minimum is found by having as little as A overlapping B as possible (or P(A and B) = 0.167)
\newline

max P(A|B) = P(A and B)|P(B) = 0.5/0.6667 = 0.75
\newline
min P(A|B) = P(A and B)|P(B) = 0.167/0.667 = 0.25

##4. Can't Please Everyone! 
Among Berkeley students who have completed w203, 3=4 like statistics. Among Berkeley students who have not completed w203, only 1=4 like statistics. Assume that only 1 out of 100 Berkeley students completes w203. Given that a Berkeley student likes statistics, what is the probability that they have completed w203?
\newline

Setup
\newline
P(like|203) = 0.75
\newline
P(like|!203) = 0.25
\newline
P(203) = 0.01
\newline
P(like) = P(like|203) * P(203) + P(like|!203) * P(!203) = (0.75)(0.01) + (0.25)(0.99) = .0075 + .2475 = .255
\newline

Answer
\newline
P(203|like) = (P(like|203) * p(203))/P(like) = .0075/(.255) = 0.029
